Probabilistic Model of Bilateral Lymphatic Spread in Head and Neck Cancer

### 2024-09-04 manuscript.qmd
make prevalences dynamic

@Fig-model-prevalences-overall shows nicely that the model is capable of accurately accounting for the most important risk factors, i.e. T-category and midline extension. As observed in the data, the model predicts that the prevalence of contralateral LNL II involvement increases from below 8% <!--  TODO: make this dynamic --> for early T-category lateralized tumors to almost 40% <!-- TODO: make this dynamic --> when the tumor is of advanced T-category and crosses the mid-sagittal line. Similarly, the prevalence of contralateral LNL III involvement increases from around 2% <!--  TODO: make this dynamic --> for early T-category lateralized tumors to almost 15% <!-- TODO: make this dynamic --> when the tumor is of advanced T-category and crosses the mid-sagittal line.

@@

The model's performance w.r.t. its ability to describe the data on contralateral nodal involvement was evaluated based on a dataset of 833<!-- TODO: make this dynamic --> patients from four institutions. To the best of our knowledge, this is the first time lymphatic tumor progression has been modelled in such detail and to such an extent. While there was work on regional tumor spread, these models were conceptually different, more limited in the LNLs for which they make predictions, and not trained with real patient data [@benson_markov_2006;@jung_development_2017].

>>>>

@Fig-model-prevalences-overall shows nicely that the model is capable of accurately accounting for the most important risk factors, i.e. T-category and midline extension. As observed in the data, the model predicts that the prevalence of contralateral LNL II involvement increases from {{< var prevalence.early_noext_iN0_cII.predicted >}}% for early T-category lateralized tumors to {{< var prevalence.late_ext_iN0_cII.predicted >}}% when the tumor is of advanced T-category and crosses the mid-sagittal line. Similarly, the prevalence of contralateral LNL III involvement increases from around {{< var prevalence.early_noext_iN0_cIII.predicted >}}% for early T-category lateralized tumors to almost {{< var prevalence.late_ext_iN0_cIII.predicted >}}% when the tumor is of advanced T-category and crosses the mid-sagittal line.

@@

The model's performance w.r.t. its ability to describe the data on contralateral nodal involvement was evaluated based on a dataset of {{< var data.num_patients >}} patients from four institutions. To the best of our knowledge, this is the first time lymphatic tumor progression has been modelled in such detail and to such an extent. While there was work on regional tumor spread, these models were conceptually different, more limited in the LNLs for which they make predictions, and not trained with real patient data [@benson_markov_2006;@jung_development_2017].
